---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  dev = "svglite"
)

pkgload::load_all(".")
```

# `gcor` package for R

<!-- badges: start -->
<!-- badges: end -->

**gcor** is an R package which provides tools for multivariate data analysis based on a  generalized correlation measure.

It features a generalized version of (absolute) correlation coefficient for arbitrary types of data, including both numerical and categorical variables. Missing values can also be handled naturally by treating them as observations of a categorical value `NA`.


## Installation

### CRAN

``` r
install.packages("gcor")
```

### GitHub (development version)

``` r
# install.packages("pak")
pak::pak("r-suzuki/gcor-r")
```

or

``` r
# install.packages("devtools")
devtools::install_github("r-suzuki/gcor-r")
```

## Documentation

### Method Overview
- [English](https://r-suzuki.github.io/gcor/method.html)
- [Japanese](https://r-suzuki.github.io/gcor/method_ja.html)


## Examples

```{r example}
library(gcor)
```

**Generalized correlation measure** takes values in $[0,1]$, which can capture
both linear and nonlinear relations.

When the joint distribution of $(X,Y)$ is bivariate normal, its theoretical value
coincides with the absolute value of the correlation coefficient.

```{r example_iris_gcor}
# Generalized correlation measure
gcor(iris)
```

The **directed generalized correlation** is another variation of the generalized correlation.
It also takes values in $[0,1]$, reaching $1$ when $Y$ is completely dependent
on $X$ (i.e., when the conditional distribution $f(Y \mid X)$ is a one-point distribution)
and $0$ when $X$ and $Y$ are independent.

```{r example_iris_dgcor}
# Dependency of Species on other variables
dgc <- dgcor(Species ~ ., data = iris)
dotchart(sort(dgc), main = "Dependency of Species")
```

With $r_g$ as the generalized correlation between $X$ and $Y$, we can define
a dissimilarity measure:

$$
d(X,Y) = \sqrt{1 - r^2_g}
$$

It can be applied to cluster analysis:

```{r example_iris_hclust}
# Clustering
gd <- gdis(iris)
hc <- hclust(gd, method = "ward.D2")
plot(hc)
```

Multidimensional scaling would serve as a good example of an application:

```{r example_iris_cmdscale}
# Multidimensional scaling
mds <- cmdscale(gd, k = 2)
plot(mds, type = "n", xlab = "", ylab = "", asp = 1, axes = FALSE,
     main = "cmdscale with gdis(iris)")
text(mds[,1], mds[,2], rownames(mds))
```
