---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  dev = "svglite"
)

pkgload::load_all(".")
```

# mdep

<!-- badges: start -->
<!-- badges: end -->

**mdep** provides measures based on mutual dependency, and other related methods for data analysis.

It features a generalized version of (absolute) correlation coefficient for arbitrary types of data, including categorical variables.

**Note that this project is in an early stage of development, so changes may occur frequently.**

## Installation

You can install the development version of mdep from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("r-suzuki/mdep")
```

## Example

```{r example}
library(mdep)
```

**Generalized correlation measure** takes values in $[0,1]$, which can capture
both linear and nonlinear relations.

When the joint distribution of $x,y$ is bivariate normal, its theoretical value
coincides with the absolute value of the correlation coefficient.

```{r example_iris_gcor}
# Generalized correlation measure
gcor(iris)
```

With $|r|$ as the generalized correlation between $x$ an $y$, we can define
a dissimilarity measure $d(x,y) = \sqrt{1 - |r|^2}$.
It can be applied to cluster analysis:

```{r example_iris_hclust}
# Clustering
gd <- gdis(iris)
hc <- hclust(gd, method = "ward.D2")
plot(hc)
```

```{r example_iris_cmdscale}
# Multidimensional scaling
mds <- cmdscale(gd, k = 2)
plot(mds, type = "n", xlab = "", ylab = "", asp = 1, axes = FALSE,
     main = "cmdscale with gdis(iris)")
text(mds[,1], mds[,2], rownames(mds))
```


The **predictability score** is another variation of the generalized correlation.
It also takes values in $[0,1]$, reaching $1$ when $y$ is perfectly predictable
by $x$ (i.e., when the conditional distribution $f(y \mid x)$ is a one-point distribution)
and $0$ when $x$ and $y$ are independent.

```{r example_iris_pscore}
# Predictability of Species from other variables
ps <- pscore(Species ~ ., data = iris)
dotchart(sort(ps), xlim = c(0, 1), main = "Predictability of Species")
```
